### 2023-06-07 frontend/merci.tex
First set-up (biblio, symbols)

%!TEX root = ../thesis_main.tex
%!TEX encoding = UTF-8 Unicode

Thank you, thank you, far too kind