### 2023-12-21 appendices/belgium_2020.tex

The Belgian whole-energy system of 2020 was largely based (88.6\% of the primary energy mix) on \og conventional fuels\fg (\ie oil and oil products (38.2\%), natural gas (29.5\%), uranium (16.3\%) and solid fossil fuels (4.6\%) while the rest mainly accounts for 26.7\,TWh of lignocellulosic and wet biomass, 12.8\,TWh of wind and 5.1\,TWh of solar \cite{spf_economy_2022}. Given the data available in the literature (mostly for the power sector) and, when not available, following the assumptions made by \citet{Limpens2020}, \Cref{tab:Belgium_2020} gives the major technologies used in 2020 to supply the different demands of \Cref{fig:cs_demands}.

\begin{table}[htbp!]
\caption{Major technologies used to supply the 2020-demands of \Cref{fig:cs_demands} in terms of share of production and installed capacity. $^{(a)}$ The decentralised heating units provide 98\% of the low-temperature heat demand. $^{(b)}$ The private mobility accounts for 80\% of the passengers mobility.}
\label{tab:Belgium_2020}
\begin{minipage}{\linewidth}
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{l c c c}
\toprule
\multirow{2}{*}{\textbf{End-use demand}} & \textbf{Major} & \textbf{Share of} & \textbf{Installed}\\
    &	 \textbf{technologies} 	& \textbf{supply} & \textbf{capacity}	 \\ 	
\midrule							
\multirow{3}{*}{Electricity}
 & Nuclear & 39\% & 5.9\,GW\\
 & CCGT & 21\% & 3.9\,GW\\
 & Wind turbines & 14\% & 5.0\,GW\\
\midrule
\multirow{3}{*}{Heat High-Temp.}
 & Gas boiler & 36\% & 3.3\,GW\\
 & Coal boiler & 30\% & 2.3\,GW\\
 & Oil boiler & 20\% & 1.5\,GW\\
\midrule
\multirow{3}{*}{Heat Low-Temp. (DEC)$^{(a)}$}
 & Oil boiler & 48\% & 21.4\,GW\\
 & Gas boiler & 40\% & 17.5\,GW\\
 & Wood boiler & 10\% & 4.4\,GW\\
\midrule
\multirow{3}{*}{Heat Low-Temp. (DHN)}
 & Gas CHP & 59\% & 0.3\,GW\\
 & Gas boiler & 15\% & 0.3\,GW\\
 & Waste CHP & 15\% & 0.1\,GW\\
\midrule
\multirow{2}{*}{Private mobility$^{(b)}$}
 & Diesel car & 49\% & 93.5\,Mpass.-km/h\\
 & Gasoline car & 49\% & 94.7\,Mpass.-km/h\\
 & HEV & 2\% & 5.9\,Mpass.-km/h\\
\midrule
\multirow{3}{*}{Public mobility}
 & Diesel bus & 43\% & 3.6\,Mpass.-km/h\\
 & Train & 43\% & 3.9\,Mpass.-km/h\\
 & CNG bus & 5\% & 0.8\,Mpass.-km/h\\
\midrule
\multirow{3}{*}{Freight mobility}
 & Diesel truck & 74\% & 62.7\,Mt.-km/h\\
 & Diesel boat & 15\% & 10.8\,Mt.-km/h\\
 & Train & 11\% & 2.5\,Mt.-km/h\\
\midrule
HVC & Naphtha/LPG cracking & 100\% & 4.6\,GW\\
Ammonia & Haber-Bosch & 100\% & 1\,GW\\
Methanol & Import & 100\% & -\\
\bottomrule							
\end{tabular}}
\end{minipage}
\end{table}