### 2024-04-05 options/symbols_and_cmd.tex

%!TEX encoding = UTF-8 Unicode
%!TEX root = ../thesis_main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	MATH SYMBOLS							%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\glsxtrsetgrouptitle{Abreviation}{Abreviation}
\glsxtrsetgrouptitle{Notation}{Notation}

% \newcommand{\sdot}{{\bullet}}
\newcommand{\sdot}{(\cdot)}

% ---------------------------------------------------------------------------- %
% ACRONYMS
%% A
\newabbreviation[group={Acronyms}]{API}{API}{Application Programming Interface}

%% B
\newabbreviation[group={Acronyms}]{BECCS}{BECCS}{Bioenergy with Carbon Capture and Storage}
\newabbreviation[group={Acronyms}]{BEMS}{BEMS}{Building Energy Management System}
\newabbreviation[group={Acronyms}]{BEV}{BEV}{Battery Electric Vehicle}
\newabbreviation[group={Acronyms}]{BTX}{BTX}{Benzene, Toluene and Xylene}

%% C
\newabbreviation[group={Acronyms}]{CAPEX}{CAPEX}{Capital Expenditure}
\newabbreviation[group={Acronyms}]{CCGT}{CCGT}{Combined Cycle Gas Turbine}
\newabbreviation[group={Acronyms}]{CCS}{CCS}{Carbon Capture and Storage}
\newabbreviation[group={Acronyms}]{CHP}{CHP}{Combined Heat and Power}
\newabbreviation[group={Acronyms}]{CNG}{CNG}{Compressed Natural Gas}

%% D
\newabbreviation[group={Acronyms}]{DC}{DC}{Direct Current}
\newabbreviation[group={Acronyms}]{DHN}{DHN}{District Heating Network}
\newabbreviation[group={Acronyms}]{DNN}{DNN}{Deep Neural Network}
\newabbreviation[group={Acronyms}]{DRL}{DRL}{Deep Reinforcement Learning}

%% E
\newabbreviation[group={Acronyms}]{ESOMs}{ESOMs}{Energy System Optimisation Models}
\newabbreviation[group={Acronyms}]{EUD}{EUD}{End-Use Demand}

%% F
\newabbreviation[group={Acronyms}]{FC}{FC}{Fuel Cell}
\newabbreviation[group={Acronyms}]{FEC}{FEC}{Final Energy Consumed}

%% G
\newabbreviation[group={Acronyms}]{GDP}{GDP}{Gross Domestic Product}
\newabbreviation[group={Acronyms}]{GHG}{GHG}{Greenhouse Gases}
\newabbreviation[group={Acronyms}]{GSA}{GSA}{Global Sensitivity Analysis}
\newabbreviation[group={Acronyms}]{GWP}{GWP}{Global Warming Potential}

%% H
\newabbreviation[group={Acronyms}]{HP}{HP}{Heat Pump}
\newabbreviation[group={Acronyms}]{HT}{HT}{High-Temperature}
\newabbreviation[group={Acronyms}]{HVC}{HVC}{High-Value Chemicals}

%% I
\newabbreviation[group={Acronyms}]{IAMs}{IAMs}{Integrated Assessment Models}
\newabbreviation[group={Acronyms}]{ICE}{ICE}{Internal Combustion Engine}
\newabbreviation[group={Acronyms}]{IEA}{IEA}{International Energy Agency}
\newabbreviation[group={Acronyms}]{IPCC}{IPCC}{Intergovernmental Panel on Climate Change}
\newabbreviation[group={Acronyms}]{IQR}{IQR}{Interquatile Range}

%% L
\newabbreviation[group={Acronyms}]{LCA}{LCA}{Life Cycle Assessment}
\newabbreviation[group={Acronyms}]{LCOE}{LCOE}{Levelised Cost of Energy}
\newabbreviation[group={Acronyms}]{LFO}{LFO}{Light Fuel Oil}
\newabbreviation[group={Acronyms}]{LOO}{LOO}{Leave-One-Out}
\newabbreviation[group={Acronyms}]{LP}{LP}{Linear Programming}
\newabbreviation[group={Acronyms}]{LPG}{LPG}{Liquefied Petroleum Gas}
\newabbreviation[group={Acronyms}]{LT}{LT}{Low-Temperature}

%% M
\newabbreviation[group={Acronyms}]{MDP}{MDP}{Markov Decision Process}
\newabbreviation[group={Acronyms}]{MMSA}{MMSA}{Methanol Market Services Asia}
\newabbreviation[group={Acronyms}]{MTBE}{MTBE}{Methyl Tert-butyl Ether}
\newabbreviation[group={Acronyms}]{MTO}{MTO}{Methanol-to-olefins}

%% N
\newabbreviation[group={Acronyms}]{NED}{NED}{Non-energy Demand}
\newabbreviation[group={Acronyms}]{NG}{NG}{Fossil Gas}
\newabbreviation[group={Acronyms}]{NN}{NN}{Neural Network}
\newabbreviation[group={Acronyms}]{NRE}{NRE}{Non-renewable Energy}
\newabbreviation[group={Acronyms}]{NSC}{NSC}{Naphtha Steam Cracker}

%% O
\newabbreviation[group={Acronyms}]{OPEX}{OPEX}{Operational Expenditure}

%% P
\newabbreviation[group={Acronyms}]{PC}{PC}{Principal Component}
\newabbreviation[group={Acronyms}]{PCs}{PCs}{Principal Components}
\newabbreviation[group={Acronyms}]{PCA}{PCA}{Principal Component Analysis}
\newabbreviation[group={Acronyms}]{PCE}{PCE}{Polynomial Chaos Expansion}
\newabbreviation[group={Acronyms}]{PDF}{PDF}{Probability Density Function}
\newabbreviation[group={Acronyms}]{PV}{PV}{Photovoltaic}

%% R
\newabbreviation[group={Acronyms}]{RE}{RE}{Renewable Energy}
\newabbreviation[group={Acronyms}]{RL}{RL}{Reinforcement Learning}

%% S
\newabbreviation[group={Acronyms}]{SAC}{SAC}{Soft Actor Critic}
\newabbreviation[group={Acronyms}]{SMR}{SMR}{Small Modular Reactor}
\newabbreviation[group={Acronyms}]{SVD}{SVD}{Singular Value Decomposition}

%% S
\newabbreviation[group={Acronyms}]{TRL}{TRL}{Technology Readiness Level}

%% U
\newabbreviation[group={Acronyms}]{UQ}{UQ}{Uncertainty Quantification}

%% V
\newabbreviation[group={Acronyms}]{VRES}{VRES}{Variable Renewable Energy Sources}

% ---------------------------------------------------------------------------- %
% ROMAN
%% A
%\glsxtrnewsymbol[group={rs},description={ith ambient particle}]{ai}{\ensuremath{A_i}}

% ---------------------------------------------------------------------------- %
% GREEK
%\glsxtrnewsymbol[group={gs},description={Yaw angle [$\deg$]}]{beta}{\ensuremath{\beta}}

% ---------------------------------------------------------------------------- %
% Sub- Superscripts
%% A
%\glsxtrnewsymbol[group={ss},description={Relative to the ambient}]{uA}{\ensuremath{\sdot_{A}}}

\renewcommand\glstreegroupheaderfmt[1]{\begingroup\textbf{#1}\vspace{-.2cm}\par\endgroup}
\glsfindwidesttoplevelname
\glsxtrsetgrouptitle{Acronyms}{Acronyms}
\glsxtrsetgrouptitle{rs}{Roman Symbols}
\glsxtrsetgrouptitle{gs}{Greek Symbols}
\glsxtrsetgrouptitle{ss}{Sub- Superscripts}
\glsxtrsetgrouptitle{o}{Operators}

%------------------------
%% A
%\newcommand{\ABL}[2]{$U_\mABL=\SIms{#1}$ - $TI_\mABL={#2}\%$}
%------------------------
%% B
\newcommand*{\boldcheckmark}{%
  \textpdfrender{
    TextRenderingMode=FillStroke,
    LineWidth=.8pt, % half of the line width is outside the normal glyph
  }{\checkmark}%
}

%------------------------
%% C

%------------------------
%% D
\newcommand{\degree}{^\circ}
\newcommand{\dd}{\mathrm{d}}

%------------------------
%% E
\newcommand{\etal}{\textit{et~al.}}

%------------------------
%% F

%------------------------
%% G

%------------------------
%% H

%------------------------
%% I

%------------------------
%% K

%------------------------
%% L

%------------------------
%% M

%------------------------
%% N

%------------------------
%% O

%------------------------
%% P
\newcommand{\myparagraph}[1]{{\vspace{0cm}\textbf{#1}\vspace{-0.3cm}}}

%------------------------
%% Q

%------------------------
%% R

%------------------------
%% S

%------------------------
%% T

%------------------------
%% U
\newcommand{\uh}{\hat{u}}
\newcommand{\uv}{\mathbf{u}}

\newcommand{\ub}{\boldsymbol{u}}

\newcommand{\uconvw}{\tilde{U}_W}

%------------------------
%% V
\newcommand{\vv}{\mathbf{v}}
\newcommand{\vb}{\boldsymbol{v}}

%------------------------
%% W
\newcommand{\wv}{\mathbf{w}}
\newcommand{\wb}{\boldsymbol{w}}
\newcommand{\WT}{{\scriptscriptstyle \mathrm{WT}}}

%------------------------
%% X
\newcommand{\xv}{\mathbf{x}}
\newcommand{\xb}{\boldsymbol{x}}
\newcommand{\xmark}{\ding{55}}

%------------------------
%% Y
\newcommand{\yv}{\mathbf{y}}
\newcommand{\yb}{\boldsymbol{y}}

%------------------------
%% Z
\newcommand{\zv}{\mathbf{z}}
\newcommand{\zb}{\boldsymbol{z}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	OPERATORS 								%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\der}[2]{\frac{d #1}{d #2}}
\newcommand{\pder}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\pdern}[3]{\frac{\partial^#3 #1}{\partial #2^#3}}

\newcommand{\avgt}[1]{\left\langle {#1}\right\rangle}
\newcommand{\avgs}[1]{\overline{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	REFERENCES 							%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\eqqref}[1]{Eq.~\eqref{#1}}
\newcommand{\seqqref}[1]{Eqs.~\eqref{#1}}
\newcommand{\eqqrefs}[2]{Eqs.~\eqref{#1} and \eqref{#2}}
\newcommand{\eqqrefc}[2]{Eqs.~\eqref{#1}, \eqref{#2}}
\newcommand{\eq}[1]{Eq.~\eqref{#1}}
% \newcommand{\eqs}[1]{Eqs.~(\ref{#1})}

\newcommand{\fig}{Fig.~\ref}
\newcommand{\figs}{Figs.~\ref}
\newcommand{\tab}{Table~\ref}
\newcommand{\sect}{Section~\ref}
\newcommand{\chap}{Chapter~\ref}
\newcommand{\chaps}{Chapters~\ref}
\newcommand{\app}{Appendix~\ref}
\newcommand{\apps}{Appendices~\ref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	CAPTIONS							%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Color palette

\usepackage{xcolor}
\definecolor{myBlue}{rgb}{0.25, 0.25, 0.75}
\definecolor{myRed}{rgb}{0.8, 0.1, 0.3}
\definecolor{myGrey0}{rgb}{0.25, 0.25, 0.25}
\definecolor{myGrey1}{rgb}{0.50, 0.50, 0.50}
\definecolor{myGrey2}{rgb}{0.75, 0.75, 0.75}

\definecolor{Blue0}{rgb}{0.7752402921953095, 0.8583006535947711, 0.9368242983467897}
\definecolor{Blue1}{rgb}{0.41708573625528644, 0.6806305267204922, 0.8382314494425221}
\definecolor{Blue2}{rgb}{0.1271049596309112, 0.4401845444059977, 0.7074971164936563}
\definecolor{Blue3}{rgb}{0.03137254901960784, 0.18823529411764706, 0.4196078431372549}

\definecolor{BluePy}{rgb}{0.69140625, 0.8164, 0.8868}
\definecolor{RedC}{rgb}{0.9453, 0.7266, 0.6602}

\definecolor{gw}{rgb}{1, 0, 0}
\definecolor{jb}{rgb}{0, 1, 0}
\definecolor{tg}{rgb}{0, 0, 1}
\definecolor{jh}{rgb}{1, 1, 0}

\newcommand{\gw}[1]{{#1}}
% \newcoxmmand{\gw}[1]{\textcolor{gw}{#1}}
\newcommand{\jb}[1]{\textcolor{jb}{#1}}
\newcommand{\tg}[1]{\textcolor{tg}{#1}}
\newcommand{\jh}[1]{\textcolor{jh}{#1}}

% Boxes at the beginning of the chapter
\usepackage{enumitem}
\definecolor{dark_blue}{RGB}{0,33,78}
\definecolor{very_light_blue}{RGB}{227, 241, 255}
\usepackage{tcolorbox} % to create boxes in the text
\newtcolorbox{mybox}[1]{colback=very_light_blue,colframe=dark_blue,fonttitle=\bfseries,title=#1}